Effects of Alternate Wetting and Drying Irrigation Technology and Nitrogen Rates on Sheath Blight of Rice

### 2020-08-12 analysis/paper/paper.Rmd
Edits to highlights. Spell out AWD

---
title: "Effects of Alternate Wetting and Drying Irrigation Technology and Nitrogen Rates on Sheath Blight of Rice"
author:
  - Adam H. Sparks:
      email: Adam.Sparks@usq.edu.au
      correspondence: true
      institute: [IRRI, USQ]
  - name: Nancy P. Castilla
    institute: IRRI
    email: n.castilla@irri.org
    correspondence: false
  - name: B. Ole Sander
    institute: IRRI
    email: b.sander@irri.org
    correspondence: false
institute:
  - IRRI: International Rice Research Institute, Los Baños, Laguna, Philippines, 4031
  - USQ: CURRENT ADDRESS -- University of Southern Queensland, Centre for Crop Health, Toowoomba, Qld 4350, Australia
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::word_document2:
    toc: no
    fig_caption: yes
    reference_docx: "../templates/template.docx"
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
csl: "../templates/plant_disease.csl" # Insert path for the bib-style
bibliography: references.bib
abstract: |
  Water and nitrogen management play vital roles in rice production. However, the mismanagement of these two crop inputs may trigger the development of sheath blight of rice, caused by *Thanatephorus cucumeris* (strain AG1-IA), which is favored by wet conditions, high relative humidity, and high nitrogen fertilizer levels. To understand how different combinations of water and nitrogen management affect sheath blight epidemics, we conducted two separate split-plot experiments with a water-saving (alternate wetting and drying) regime and a traditional puddled regime combined with differing nitrogen treatments in the dry seasons of 2015 and 2016. Disease was scored using the same methodology in both experiments using a sheath blight assessment scale for field evaluation developed at the International Rice Research Institute to assess the severity on infected sheaths and leaves while sheath blight incidence on tillers was counted per hill. We were unable to detect any clear differences in the incidence or severity of tiller sheath blight due to irrigation treatment. Leaf severity for the AWD treatment was clearly lower than traditional puddling in the 2015 trial but we were unable to detect any clear differences in 2016. However, the level of disease severity remained low in both years, < 1&nbsp;%. Our findings indicate that farmers can adopt water-saving technologies without risking increased sheath blight incidence. We suggest that further cross-disciplinary research in this area is warranted.
keywords: |
  Rice; Water Management; Nitrogen Management; Sheath Blight; *Thanatephorus cucumeris*
highlights: |
  * The effects of alternate wetting and drying and nitrogen fertilisation rates on sheath blight disease incidence and severity were investigated.
  * Increasing rates of nitrogen led to increased levels of sheath blight incidence and sheath blight severity.
  * AWD had no detectable effect on either sheath blight incidence or sheath blight severity.
  * The adoption of AWD technologies does not appear to lead to increased sheath blight incidence or sheath blight severity.
---

>>>>

---
title: "Effects of Alternate Wetting and Drying Irrigation Technology and Nitrogen Rates on Sheath Blight of Rice"
author:
  - Adam H. Sparks:
      email: Adam.Sparks@usq.edu.au
      correspondence: true
      institute: [IRRI, USQ]
  - name: Nancy P. Castilla
    institute: IRRI
    email: n.castilla@irri.org
    correspondence: false
  - name: B. Ole Sander
    institute: IRRI
    email: b.sander@irri.org
    correspondence: false
institute:
  - IRRI: International Rice Research Institute, Los Baños, Laguna, Philippines, 4031
  - USQ: CURRENT ADDRESS -- University of Southern Queensland, Centre for Crop Health, Toowoomba, Qld 4350, Australia
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::word_document2:
    toc: no
    fig_caption: yes
    reference_docx: "../templates/template.docx"
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
csl: "../templates/plant_disease.csl" # Insert path for the bib-style
bibliography: references.bib
abstract: |
  Water and nitrogen management play vital roles in rice production. However, the mismanagement of these two crop inputs may trigger the development of sheath blight of rice, caused by *Thanatephorus cucumeris* (strain AG1-IA), which is favored by wet conditions, high relative humidity, and high nitrogen fertilizer levels. To understand how different combinations of water and nitrogen management affect sheath blight epidemics, we conducted two separate split-plot experiments with a water-saving (alternate wetting and drying) regime and a traditional puddled regime combined with differing nitrogen treatments in the dry seasons of 2015 and 2016. Disease was scored using the same methodology in both experiments using a sheath blight assessment scale for field evaluation developed at the International Rice Research Institute to assess the severity on infected sheaths and leaves while sheath blight incidence on tillers was counted per hill. We were unable to detect any clear differences in the incidence or severity of tiller sheath blight due to irrigation treatment. Leaf severity for the AWD treatment was clearly lower than traditional puddling in the 2015 trial but we were unable to detect any clear differences in 2016. However, the level of disease severity remained low in both years, < 1&nbsp;%. Our findings indicate that farmers can adopt water-saving technologies without risking increased sheath blight incidence. We suggest that further cross-disciplinary research in this area is warranted.
keywords: |
  Rice; Water Management; Nitrogen Management; Sheath Blight; *Thanatephorus cucumeris*
highlights: |
  * The effects of alternate wetting and drying and nitrogen fertilisation rates on sheath blight disease incidence and severity were investigated.
  * Increasing rates of nitrogen led to increased levels of sheath blight incidence and sheath blight severity.
  * Alternate wetting and drying had no detectable effect on either sheath blight incidence or sheath blight severity.
  * The adoption of alternate wetting and drying technology does not appear to lead to increased sheath blight incidence or sheath blight severity.
---
